### 2017-05-11 paper2.tex
Add files via upload

The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage. Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}). Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input. In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible and away from latent. These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI). To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology. The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions. Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales \cite{Peng2012,Zhao2014}. However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:

\begin{enumerate}
	\item Geometric undersampling of 3-dimensional surface structure
	\item Satellite overpass cycles and sensor sampling regimes that result in temporal discontinuity of the surface temperature time series
	\item Clear-sky bias that contribute to an overestimation of "all-sky" surface temperatures
\end{enumerate}

@@

Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban T\textsubscript{surf} measurements are representative of coherent urban patches and free from confounding influences (i.e. atmospheric and emissivity effects). Meta-analysis of the air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}. Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis. In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}. Using the method and parameterization described in Chapter \ref{paper1} twl long-term climatologies of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) were derived from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer. The method, and an associated parameterization scheme, was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. Geometric sampling biases are a result of lens distortions and the sensor cosine response. However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface. The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} urban T\textsubscript{surf} under all-sky conditions for sUHI analysis. These measurements are often made as a part of the net radiation determination for urban energy balance studies.

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing longwave radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry. This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 to 12 \si{\kelvin} \citep{Voogt1995,Lagouarde2012}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

@@

Both geometric and temporal shortcomings limit the representivity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI. The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective. These biases lead us to the question: What is the nature of the true sUHI? How does it vary temporally and over a full range of weather conditions? To address these questions, this study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}. These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.

@@

This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive a time-continuous eight month climatology of urban T\textsubscript{hem, r} for sUHI analysis. A thorough discussion of the method is included in Chapter \ref{paper1}.

@@

Remote sensing of TIR is subject to atmospheric influence from gaseous and aerosol absorbers between the surface and the sensor. A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal. These effects can result in differences of up to 8 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR. Atmospheric effects are particularly important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence. This is particularly salient for near-ground, broadband radiometers, as band by band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}. As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance. Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response. For each time step, the method generates a LUT of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity. A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing. The work flow, described in Figure \ref{flow}, is repeated at 30-minute intervals over the eight month study period to retrieve a climatology of urban T\textsubscript{hem, r} for sUHI analysis.

First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}. Using a digital building model (DBM), SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

Second, version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} is used to model spectral at-sensor radiances for each angular path length. Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step. Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal. Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers transmit radiation at much smaller wavenumbers (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).

@@

Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 16-day period. In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step. Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.

Sensor placement sensitivity testing in section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights. This introduces a potential bias in T\textsubscript{hem, r} as changing sensor position relative to canyon features alters view factor proportions of rooftops, roads, and walls. The sensitivity of T\textsubscript{hem, r} to sensor placement decreases with height and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}. As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than normal. Although pyrgeometers at both the Sperrstrasse and are mounted below the height at which T\textsubscript{hem,r} is spatially invariant, sensitivity tests in section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement at the Sperrstrasse canyon. Similar or lesser biases as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}). In addition, T\textsubscript{hem, r} shows good performance relative to T\textsubscript{plan} when compared to T\textsubscript{comp}, particularly under clear sky conditions when microscale spatiotemporal variations in urban T\textsubscript{surf} and sUHI magnitudes are greatest. As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small.

@@

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada. Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign. Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site. Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.

@@

To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme. EPiCC urban and rural data were collected from an urban/suburban site in central Vancouver ("Sunset Tower") and from a rural site approximately 25 \si{\kilo\meter} to the SSW ("Westham Island"). These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.

@@

sUHI and clhUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,

@@

To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP. Observations at Village Neuf were intermittent and often discontinuous As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI. As sUHI develop, preference was given to days with continuous observations from early afternoon through the night. For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals to compare patterns of sUHI development for the two sites. Mean normalized sUHI was then calculated at each time step over the nine day period. A normalized T\textsubscript{hem, b} was also calculated (using maximum T\textsubscript{hem, b} for each day) to compare T\textsubscript{surf} regimes between the two sites.

@@

Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHi\textsubscript{norm} for the BUBBLE campaign. By restricting UHI values to a range between one and zero, diurnal patterns of sUHI and clUHI can be compared directly. Comparison of the two in summer shows a peak sUHI approximately two hours after solar noon (when clUHI is small) and peak clUHI approximately two hours after sunset (when sUHI is small). Values near zero in the winder months indicate little variation in both sUHI and clUHI.

@@

Figure \ref{vanc_suhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the EPiCC campaign. The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but does not persist as far into the late evening hours.

@@

Figure \ref{bxsUHI_season_bubble} shows hemispherical sUHI magnitudes for the eight month BUBBLE campaign and binned for winter (December through February) and summer (April through July) seasons. Diurnal variability in sUHI magnitudes is concentrated in summer months, as well as seasonal sUHI\textsubscript{max}.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP. Each plot includes mean sUHI, as well as case days representing sUHI under clear sky and clouded sky conditions.}
	\label{bx_suhi_compare}
\end{figure}

@@

Figure \ref{meteo_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign. Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}.

>>>>

The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage. Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}). Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; and traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input. In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible heat and away from latent heat. These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI). To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology. The proliferation of satellite and aerial thermal remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions. Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island patterns globally at large spatial scales \citep{Peng2012,Zhao2014}. However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:

\begin{enumerate}
	\item Geometric undersampling of 3-dimensional surface structure
	\item Satellite overpass cycles and sensor sampling regimes that result in temporal discontinuity of the surface temperature time series
	\item Clear-sky bias that contributes to an overestimation of "all-sky" surface temperatures
\end{enumerate}

@@

Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban T\textsubscript{surf} measurements are representative of coherent urban patches and free from confounding influences (i.e. atmospheric and emissivity effects). Meta-analysis of the aUHI literature shows that these goals are rarely satisfied \citep{Stewart2011}. Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis. In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}.

%Using the method and parameterization described in Chapter \ref{paper1} two long-term climatologies of hemispherical radiometric urban T\textsubscript{surf} were derived from time continuous near-ground upwelling TIR measured from an inverted pyrgeometer. These climatologies represent the first time-continuous seasonal analysis of the urban effect on T\textsubscript{surf}. The method, and an associated parameterization scheme, was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. Geometric sampling biases are a result of lens distortions and the sensor cosine response. However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface. The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} urban T\textsubscript{surf} under all-sky conditions for sUHI analysis. These measurements are often made as a part of the net radiation determination for urban energy balance studies.

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing longwave radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-FOV TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry. This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 to 12 \si{\kelvin} \citep{Voogt1995,Lagouarde2012}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

@@

Both geometric and temporal shortcomings limit the representivity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI. The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective. These biases lead us to the question: What is the nature of the true sUHI? How does it vary temporally and over a full range of weather conditions? To address these questions, this study presents the first time-continuous, climatological analysis of sUHI derived from geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. Geometric sampling biases are a result of lens distortions and the sensor cosine response. However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface. The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}. sUHI calculated for the eight month BUBBLE campaign in Basel, Switzerland and a year long subset of the EPiCC campaign in Vancouver, Canada are used to analyze the diurnal and seasonal character of sUHI and to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.

@@

This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive time-continuous urban T\textsubscript{hem, r} for sUHI analysis. A thorough discussion of the method is included in Chapter \ref{paper1}.

@@

Remote sensing of TIR is subject to atmospheric influence from gaseous and aerosol absorbers between the surface and the sensor. A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal. These effects can result in differences of up to 10 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR. Atmospheric effects are important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence and degrade comparability. This is particularly salient for near-ground, broadband radiometers, as band-by-band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}. As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance. Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple LOS correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response. For each time step, the method generates a LUT of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity. A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing. The work flow, described in Figure \ref{flow}, is repeated at 30-minute intervals over the eight month study period to retrieve a climatology of urban T\textsubscript{hem, r} for sUHI analysis.

First, surface-to-sensor path length and view factor geometries are calculated with the SUM sensor view model \citep{Soux2004}. Using a digital building model (DBM), SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

Second, version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} is used to model spectral at-sensor radiances for each angular path length. Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step. Spectral radiances are modeled over a bandpass of 1 to 2300 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal. Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2300 \si{cm^{-1}}), as silicone domed pyrgeometers transmit radiation at much smaller wavenumbers (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).

@@

Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted a robust evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 14-day period. In the evaluation, a modified version of the correction method uses T\textsubscript{hem, b} inferred from the lowest upwelling longwave measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step. Modeled and measured irradiances match well over wide range of day and night mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.

Sensor placement sensitivity testing in Section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights. This introduces a potential bias in T\textsubscript{hem, r} as changing sensor position relative to canyon features alters view factor proportions of rooftops, roads, and walls. The sensitivity of T\textsubscript{hem, r} to sensor placement decreases as sensor height is increased, and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}. As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than typical. Although pyrgeometers in both studies utilized in this campaign are mounted below the height at which T\textsubscript{hem, r} is spatially invariant, sensitivity tests in Section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement in the BUBBLE campaign. Similar biases are expected in analysis of Vancouver sUHI as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}). In addition, T\textsubscript{hem, r} shows good performance relative to T\textsubscript{plan} when compared to T\textsubscript{comp}, particularly under clear sky conditions when microscale spatiotemporal variations in urban T\textsubscript{surf} and sUHI magnitudes are greatest. As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small and relatively consistent.

@@

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in Section \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a LUT, where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance to calculate the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and a year long subset of the the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada. Using the method described in Section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign. Urban data were observed from a tower located in a mid-rise, largely residential neighborhood near city center at the "Basel Sperrstrasse" site. Rural reference data were observed from the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.

@@

To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme. EPiCC urban and rural data were collected from an urban residential neighborhood site in central Vancouver (”Sunset Tower”) and from a rural site over a flat, non-irrigated, unmanaged field approximately 25 km to the SSW (”Westham Island”) near to the coast.  These data provide a second, long term, time-continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.

@@

sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,

@@

To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP. Observations at Village Neuf were intermittent and often discontinuous As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI. As sUHI development is most variable in the late afternoon and evening hours, preference was given to days with continuous observations from early afternoon through the night. For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30 \si{\minute} intervals to compare patterns of sUHI development for the two sites. Mean normalized sUHI was then calculated at each time step over the nine day period. A normalized T\textsubscript{hem, b} was also calculated (using maximum T\textsubscript{hem, b} for each day) to compare T\textsubscript{surf} regimes between the two sites.

@@

Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHI\textsubscript{norm} for the BUBBLE campaign. By restricting UHI values to a range between one and zero, diurnal patterns of sUHI and clUHI can be compared directly. Comparison of the two in summer shows a peak sUHI approximately two hours after solar noon (when clUHI is small) and peak clUHI approximately two hours after sunset (when sUHI is small). Values near zero in the winder months indicate little variation in both sUHI and clUHI.

@@

Figure \ref{vanc_suhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the EPiCC campaign. The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but degrades more quickly in the late evening hours.

@@

Figure \ref{bxsUHI_season_bubble} shows hemispherical sUHI magnitudes for the eight month BUBBLE campaign and binned for winter (December through February) and summer (April through July) seasons. Diurnal variability in sUHI magnitudes and seasonal sUHI\textsubscript{max} is concentrated in summer months.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the BUBBLE IOP. Each plot includes mean sUHI, as well as case days representing sUHI under clear sky and clouded sky conditions.}
	\label{bx_suhi_compare}
\end{figure}

@@

\subsubsection{BUBBLE}

Figure \ref{meteo_sol} shows mean and maximum hemispherical sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign. Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}.
